\documentclass[Main.tex]{subfiles}

%these make the chapter/section numbers correct if built individually
%obviously you'd have to adjust these to be accurate.
%\setcounter{chapter}{2}
%\setcounter{section}{3}

%\externaldocument{Introduction}

\begin{document}
\ourchapter{Evaluation Plan}\label{ch:evaluation plan}

This chapter outlines the evaluation methodology to be used to assess the performance of the proposed generative model. Given that the task involves several objectives, evaluation must account for multiple dimensions. A combination of human evaluation and quantitative metrics is proposed. It should be noted that human evaluation is expected to be subject to significant bias. Evaluators are likely to be predominantly European, American, or Chinese, meaning that judgements on pronounceability, creativity, plausibility, and cultural coherence will be more reliable for cultures familiar to the evaluators and less reliable for others. In particular, evaluators may struggle to assess the authenticity of names from cultures they have no exposure to, potentially conflating cultural plausibility with mere phonetic familiarity. As a baseline, the model should be able to reconstruct as faithfully as possible the input from its latent representation in an autoregressive manner. Reconstruction quality provides an indication that the code retains sufficient information about the input sequence. Levenshtein distance \cite{levenshtein_1966} between strings of characters is the minimum number of single-character edits (insertions, deletions, substitutions) required to change one sequence into the other. It measures string similarity and, in this context, is to be used to evaluate the distance between the input and output sequences of characters. In the context of NLP, it is common to normalise the Levenshtein difference by length in order to make it length-agnostic. While being a useful baseline, being able to reconstruct the input sequence of characters is not the main goal of the model. At its core, the model should be able to generate pronounceable, plausible, culturally coherent, novel and diverse anthroponyms and toponyms from existing or fictional cultures. The following sections outline the proposed evaluation metrics for each objective.\\

\subsection{Pronounceability}
One of the key goals of this model is to generate sequences of characters that a native speaker would be able to pronounce with relative ease. For this to happen, the sequences generated by the model should align with the phonotactic rules of the target language or culture. Pronounceability is to be primarily assessed through human evaluation, where participants rate the relative ease or difficulty with which a generated sequence could be spoken aloud by a native speaker of a given language or cultural background. To complement human evaluation, additional proxy measures are used to provide a more quantitative signal. As in \cite{graves_generating_2013}, character-level language model perplexity (PPL) is to be employed to measure how likely a sequence is under a model trained on a dataset of real anthroponyms and toponyms: lower perplexity indicates greater conformity to learned character-level distributions and is therefore used as a proxy for pronounceability. In addition, the proportion of character
$n$-grams in generated names that appear in the training data \cite{jurafsky_speech_2026} may also be used as a proxy for phonotactic plausibility.\\

\subsection{Diversity and Novelty}
Another key objective is for the model to generate new and varied names, a task that can be interpreted as a creative exercise. Due to the challenges involved in measuring creativity, we propose to use a combination of human evaluation and quantitative proxy measures. In the context of name generation, creativity can be interpreted as the ability to generate names that are both diverse and novel, avoiding both repetition among generated outputs and memorisation of training examples, thereby preventing mode collapse to a limited set of outputs. Novelty is to be defined with respect to both the training data and the conditioning examples provided at inference time. With respect to the training data, novelty is measured using the normalised Levenshtein distance between the generated sequence and its nearest neighbour in the training set, where larger distances can be interpreted as a proxy for greater novelty. In the conditional setting, novelty is additionally assessed with respect to the input prompts provided at inference time, ensuring that the model does not simply replicate or trivially transform conditioning examples. Diversity, on the other hand, measures variation among generated samples and is quantified using metrics such as duplicate rate and pairwise normalised Levenshtein distance between generated sequences, where higher average distances indicate reduced mode collapse and greater output variability. Self-BLEU \cite{zhu_texygen_2018} may also be used as an additional diversity metric. This measures the similarity of each generated name to the rest of the generated set, such that lower self-BLEU can be interpreted as a proxy for greater diversity.\\

\subsection{Plausibility and Cultural Coherence with Existing Cultures}
Next, the model should be able to generate anthroponyms or toponyms that are plausible names from a given culture or language, as directly or indirectly indicated at inference time. The fact that a sequence of characters sounds as if it comes from a given language does not imply that it also sounds like an anthroponym or toponym from that language, hence this objective is clearly two-fold and can be split into name plausibility and cultural coherence. We propose to utilise a combination of human evaluation and quantitative evaluation metrics to assess these objectives. First of all, human evaluation is to be used to assess the capability of the model to generate plausible anthroponyms or toponyms from a given culture. Furthermore, two separate classifiers are trained: one on anthroponyms and one on toponyms. At evaluation time, generated names are passed through the corresponding classifier and cultural authenticity is measured as the proportion of generated names correctly classified as belonging to the target culture. As discussed above, character-level perplexity also serves as a proxy for phonetic and orthographic plausibility of generated outputs. Finally, cultural coherence is to be measured by assessing the ability of the model to organise its latent space by culture, such that names from the same culture are closer than names from different cultures. Silhouette score \cite{rousseeuw_silhouettes_1987} is to be used for this purpose, such that, if the model achieves a good Silhouette score, then one would expect that interpolated names within cultural or linguistic clusters would generate names that exhibit cultural coherence to the same culture.\\

\subsection{Generation for Fictional Cultures}
Finally, the model should be able to generate plausible anthroponyms or toponyms from fictional cultures. To assess this goal, human evaluation is to be used in addition to a number of quantitative proxies. Again, Silhouette score is to be used to measure the ability of the model to organise its latent space in a way that is semantically meaningful and, in particular, such that distance between cultural clusters is a proxy of the similarity or dissimilarity between their languages.  If the model were able to achieve a high Silhouette score, one would expect interpolated names between two cultural clusters to produce a name that sounds like it somewhat belongs to both cultures, while being neither of them. In such a case, a culture classifier trained on real names should assign the interpolated name to a culture that is intermediate between the two source cultures or flag it as unrecognised, thereby indicating its genuine novelty.\\

\end{document}